---
title: "p8105_hw1_zw2899"
author: "Ziqing Wang"
date: "2022-09-21"
output: github_document
---

```{r, echo=FALSE}
library(tidyverse)
library(ggplot2)
```
# Problem 1 
Load the penguins data set
```{r}
data("penguins", package = "palmerpenguins")
```

The penguin dataset contains `r ncol(penguins)` columns and `r nrow(penguins)` rows. The data contains three different penguine species (`r unique(penguins$species)`) from three different islands (`r unique(penguins$island)`). All data points were collected in years `r unique(penguins$year)`. `r table(penguins$sex)["female"]` of the included penguins are female, while `r table(penguins$sex)["male"]` are male. The mean flipper length in mm is `r round(mean(penguins$flipper_length_mm, na.rm=T), 2)`, ignoring missing values in this variable.

Below is a scatterplot 
```{r}
ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species)) + 
  geom_point() +
  xlab("bill length (mm)") + ylab("flipper length (mm)") +
  ggtitle("flipper length vs. bill length by species")

ggsave("flipper_vs_bill_scatter_plot.pdf", height=4, width=5)
```

# Problem 2

Create the specified data frame:

```{r}
q2_df = tibble(
  numeric_vec = rnorm(10), # a random sample of size 10 from the standard normal distribution
  logical_vec = numeric_vec > 0, # a logical vector indicating whether each sample element from above is > 0
  char_vec = c("rainy", "rainy", "snowy", "sunny", "cloudy", 
               "stormy", "foggy", "sunny", "drizzling", "windy"), # a character vector of length 10
  factor_vec = factor(c("sad", "sad", "sad", "happy", "neutral",
                        "sad", "neutral", "happy", "neutral", "neutral")) # a factor vector with 3 different levels
)

q2_df
```
The mean of the numerical vector and the logical vecctor can be computed successfully, which equal to `r mean(pull(q2_df, var=1))` and `r mean(pull(q2_df, var=2))`, respectively. However, R raises warnings when attempting to take the mean of the character vector and the factor vector and returns NAs:

```{r}
mean(pull(q2_df, var=3))
mean(pull(q2_df, var=-1))
```

Now attempt to apply the as.numeric() function to the logical, character, and factor variables;

```{r, eval=FALSE}
as.numeric(q2_df$logical_vec)
as.numeric(q2_df$char_vec)
as.numeric(q2_df$factor_vec)
```

The as.numeric function turns the FALSE to 0 and TRUE to 1 in the logical vector. Therefore, R is able to successfully calculate its mean. The as.numeric function is not able to coerce character strings to numerical values. Therefore, it returns NAs. As a result, R turns NA when trying to calculate the mean of a character vector. As for the factor vector, the as.numeric function coerces the factor levels to positive integers by alphabetical order. However, elements in the factor elements themselves are neither logical or numeric values. Therefore, R cannot calculate its mean and return NA as shown above. However, after coercing the factor vector to a numerical vector, its mean can be computed as `r mean(as.numeric(q2_df$factor_vec))`.



